Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2018-08-29 Ben_Manuscripts/structure_paper/Draft.tex
some thoughts on intro and abstract.

  Nanostructured porous membranes made from the cross-linked hexagonal phase
  of self-assembled lyotropic liquid crystals (LLCs) are a promising material
  for selective separations. In this work, we investigate an experimentally
  characterized LLC membrane using atomistic molecular modeling. In particular,
  we compare simulated X-ray diffraction (XRD) patterns with experimental XRD
  to quantify and understand the differences between simulation and experiment.
  We find that pores are likely composed of 5 columns of stacked LLC monomers
  which surround each hydrophilic core. We learned that these columns likely 
  move independently of each other over time scales which we can not simulate.
  We find that some of the structure previously attributed to monomer tail 
  tilt is likely instead due to ordered tail packing. Although this system
  has been reported as dry, we show that small amounts of water are necessary
  to fully reproduce all features from the experimental XRD pattern due to
  asymmetries introduced by hydrogen bonds between monomer head groups and
  water molecules. We explore the composition and structure of the nanopores
  and reveal that there exists a composition gradient rather than an abrupt
  partition between the hydrophilic and hydrophobic regions. 
  % BJC13: If this is to be included here, it needs some rewording. I allude to
  % this when talking about the independently moving columns above. 
  %A caveat is
  %that the time scales of the dynamics are extremely long for this system, 
  %resulting in simulated structures that appear too ordered which requires
  %careful examination of the metastable states observed in order to draw any
  %conclusions. 
  The clearer picture of the nanoscopic structure of these membranes provided
  in this study will enable a better understanding of the mechanisms of small
  molecule transport within these nanopores.

@@

  More highly selective membranes would be extremely useful for both the
  recovery of valuable dissolved species from and the purification of complex
  aqueous solutions. The supply of fresh water is being increasingly strained
  by population growth, industrialization and climate change
  \cite{elimelech_future_2011}. Consequently, people are turning towards
  treatment of unconventional water sources such as seawater and wastewater. 
  In addition to obtaining fresh water from them, these water sources, 
  particularly municipal and industrial wastewater, have been found to contain 
  valuable resources which can be separated using highly selective membranes~\cite{sales_resource_2015}.

  Reverse osmosis (RO) and nanofiltration (NF) are two prevailing membrane 
  filtration processes that can be used to separate solutes on the order of 1 
  nm in size and smaller, as well as ions. Both apply hydraulic pressure to 
  the feed solution in order to overcome osmotic pressure and force water
  and unfiltered components through the membrane. RO membranes typically have
  a thin film composite architecture which, at minimum, consists of a porous
  support layer and a dense polymer active layer. Separation occurs based on the differences in
  solubility and diffusivity of feed solution constituents with respect to the
  active layer. Nanofiltration membranes have explicit pores on the order of 1
  nm in diameter which are typically tortuous and polydisperse in size. They 
  separate primarily based on size exclusion as well as Donnan exclusion if the
  membrane has a surface charge~\cite{van_der_bruggen_review_2003}.

  The energy consumed during the purification of seawater and wastewater 
  using RO and NF membranes can be reduced by increasing their selectivity 
  towards water \cite{werber_materials_2016}. While researchers often focus on
  improving membrane permeability, the flux of water through a given membrane,
  it has been shown that increased water permeability only slightly decreases
  energy requirements \cite{cohen-tanugi_quantifying_2014}.
%  since the main energy requirement for a membrane separation is that needed
%  to create the hydraulic pressure necessary in order to overcome the 
%  osmotic pressure of the feed solution\cite{cohen-tanugi_quantifying_2014}. 
  Often times, multiple membrane passes are required in order to sufficiently
  reduce contaminant levels. For example, boric acid is toxic to humans and 
  is present in seawater at levels well above acceptable drinking standards.
  It is difficult to sufficiently separate boric acid in a single pass, meaning
  multiple passes are needed and energy requirements nearly double \cite{fritzmann_state---art_2007}.
%  This is necessary in the case where high-purity water
%  is needed \cite{singh_production_2009}, when separating small neutral solutes \cite{ozaki_rejection_2002} or
%  when removing boric acid from seawater \cite{fritzmann_state---art_2007}.  
  The most effective way to reduce energy consumption
  would be to reduce the number of membrane passes for a given system using 
  more selective membranes. 
%   More highly selective membranes have the potential
%  to reduce the number of membrane passes in addition to eliminating the need 
%  for other post-treatment steps necessitated by incomplete separations.
  %dissolved species which negatively
%  affect human health.

@@

  Industrial waste-waters are often quite complex with up to six times more 
  total dissolved solids than seawater\cite{werber_materials_2016}. For example, flowback water, produced
  during hydraulic fracturing of shale formations consists of relatively high
  concentrations of salts, metals, and soluble organic compounds. The majority
  of this water is disposed through deep well injection, however there is a growing public concern about
  its management which has prompted the use of separation technologies such as 
  RO and NF in order to reduce the volume of contaminated water~\cite{gregory_water_2011}.
  Some of the dissolved organic compounds in flowback water, such as acetate, 
  are potentially valuable and can be recovered with highly selective membranes
  ~\cite{dischinger_application_2017}.

>>>>

  Nanostructured porous membranes made from the cross-linked hexagonal phase
  of self-assembled lyotropic liquid crystals (LLCs) are a promising material
  for selective separations. In this work, we investigate an experimentally
  characterized LLC membrane using atomistic molecular modeling. In particular,
  we compare simulated X-ray diffraction (XRD) patterns with experimental XRD
  to quantify and understand the differences between simulation and experiment.
  We find that pores are likely composed of 5 columns of stacked LLC monomers
  which surround each hydrophilic core. We learned that these columns likely 
  move independently of each other over time scales which we can not simulate.
  We find that some of the structure previously attributed to monomer tail 
  tilt is likely instead due to ordered tail packing. Although this system
  has been reported as dry, we show that small amounts of water are necessary
  to fully reproduce all features from the experimental XRD pattern due to
  asymmetries introduced by hydrogen bonds between monomer head groups and
  water molecules. We explore the composition and structure of the nanopores
  and reveal that there exists a composition gradient rather than an abrupt
  partition between the hydrophilic and hydrophobic regions. 
  % BJC13: If this is to be included here, it needs some rewording. I allude to
  % this when talking about the independently moving columns above. 
  %A caveat is
  %that the time scales of the dynamics are extremely long for this system, 
  %resulting in simulated structures that appear too ordered which requires
  %careful examination of the metastable states observed in order to draw any
  %conclusions. 
  %MRS14: OK, let's talk through it.  I feel that this is an important caveat to add so
  %that people don't think we are claiming more than we are, and therefor should be mentioned
  % explicitly, rather than the implication left above.
  The clearer picture of the nanoscopic structure of these membranes provided
  in this study will enable a better understanding of the mechanisms of small
  molecule transport within these nanopores.

@@

  %MRS14: I feel this introduction spends a bit too much time talking at too general a level.  The paper introduction to the problem tackled in the paper.  One simple way to tighten it up would be to improve the connection between the thesis and the data.  You say that highly selective membranes would be useful, but then don't provide evidence that they are the solution to the problems listed. 
  More highly selective membranes would be extremely useful for both the
  recovery of valuable dissolved species from and the purification of complex
  aqueous solutions. The supply of fresh water is being increasingly strained
  by population growth, industrialization and climate change
  \cite{elimelech_future_2011}. Consequently, people are turning towards
  treatment of unconventional water sources such as seawater and wastewater. 
  In addition to obtaining fresh water from them, these water sources, 
  particularly municipal and industrial wastewater, have been found to contain 
  valuable resources which can be separated using highly selective membranes~\cite{sales_resource_2015}.

  Reverse osmosis (RO) and nanofiltration (NF) are two prevailing membrane 
  filtration processes that can be used to separate solutes on the order of 1 
  nm in size and smaller, as well as ions. Both apply hydraulic pressure to 
  the feed solution in order to overcome osmotic pressure and force water
  and unfiltered components through the membrane. RO membranes typically have
  a thin film composite architecture which, at minimum, consists of a porous
  support layer and a dense polymer active layer. Separation occurs based on the differences in
  solubility and diffusivity of feed solution constituents 
  %MRS14:
  %with respect to the
  in the 
  active layer. Nanofiltration membranes have explicit pores on the order of 1
  nm in diameter which are typically tortuous and polydisperse in size. They 
  separate primarily based on size exclusion as well as Donnan exclusion if the
  membrane has a surface charge~\cite{van_der_bruggen_review_2003}.

  %MRS14: is selectivity towards water important, or selectivity against what one wants filtered out?
  The energy consumed during the purification of seawater and wastewater 
  using RO and NF membranes can be reduced by increasing their selectivity 
  towards water \cite{werber_materials_2016}. While researchers often focus on
  improving membrane permeability, the flux of water through a given membrane,
  it has been shown that increased water permeability only slightly decreases
  energy requirements \cite{cohen-tanugi_quantifying_2014}.
%  since the main energy requirement for a membrane separation is that needed
%  to create the hydraulic pressure necessary in order to overcome the 
%  osmotic pressure of the feed solution\cite{cohen-tanugi_quantifying_2014}. 
  Often times, multiple membrane passes are required in order to sufficiently
  %MRS14: not sure the boric acid part is that relevant, since we don't really talk about 
  %that in the rest of the paper.
  reduce contaminant levels. For example, boric acid is toxic to humans and 
  is present in seawater at levels well above acceptable drinking standards.
  It is difficult to sufficiently separate boric acid in a single pass, meaning
  multiple passes are needed and energy requirements nearly double \cite{fritzmann_state---art_2007}.
%  This is necessary in the case where high-purity water
%  is needed \cite{singh_production_2009}, when separating small neutral solutes \cite{ozaki_rejection_2002} or
%  when removing boric acid from seawater \cite{fritzmann_state---art_2007}.  
  The most effective way to reduce energy consumption
  would be to reduce the number of membrane passes for a given system using 
  more selective membranes. 
%   More highly selective membranes have the potential
%  to reduce the number of membrane passes in addition to eliminating the need 
%  for other post-treatment steps necessitated by incomplete separations.
  %dissolved species which negatively
%  affect human health.

@@

  %MRS14: could possibly be food details, if selective membranes are the key to purifying industrial waste waters. 
  Industrial waste-waters are often quite complex with up to six times more 
  total dissolved solids than seawater\cite{werber_materials_2016}. For example, flowback water, produced
  during hydraulic fracturing of shale formations consists of relatively high
  concentrations of salts, metals, and soluble organic compounds. The majority
  of this water is disposed through deep well injection, however there is a growing public concern about
  its management which has prompted the use of separation technologies such as 
  RO and NF in order to reduce the volume of contaminated water~\cite{gregory_water_2011}.
  Some of the dissolved organic compounds in flowback water, such as acetate, 
  are potentially valuable and can be recovered with highly selective membranes
  ~\cite{dischinger_application_2017}.
